On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-07 main.tex
Update main.tex
typo
Update main.tex
Updated Section 5

The runtime for cases without a strait-line program in the input are as follows. In these cases a polynomial $f$ of degree $d$ is given in the input, and a strait-line program is built for it by computing all the terms. In this case, $L \geq d^n$ and $b \geq \htt(f) = s,$ so that the number of bit operations becomes
\begin{align*}
      &O^{\sim}(nk(Lb+d^{2n}(s + \log 1/\epsilon^2+d)(L+d)\log(s + \log 1/\epsilon^2)))\\
      &= 
O^{\sim}(nk(d^ns+d^{2n}(s + \log 1/\epsilon^2+d)(d^n+d)\log(s + \log 1/\epsilon^2)))\\
      &= 
O^{\sim}(nk(d^ns+(d^{3n}+d^{2n+1})(s + \log 1/\epsilon^2+d)\log(s + \log 1/\epsilon^2)))\\
     &= 
O^{\sim}(nk(d^{3n}(s + \log 1/\epsilon^2+d)\log(s + \log 1/\epsilon^2)))\\
     &= 
O^{\sim}(nk(d^{3n+1}(s + \log 1/\epsilon^2)\log(s + \log 1/\epsilon^2))),
\end{align*}
and the degree and height of the output does not change.

@@

% \section{Proof of Theorem 2.1}
% %
% Recall Theorem 3.1: \textit{Suppose that $f \in \Q[X_1,\hdots,X_n]$ satisfies \textbf{H}. Then there exists a Zariski closed set $\D \subset \mathbb{C}^{n^2}$ of degree at most $D^{\textrm{\td}}$ such that, if $\textbf{A}$ is chosen from  $\C^{n^2}-\Delta$ ( $\Q^{n^2}-\Delta$?) then $f^A$ satisfies $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$} 
% %
% \begin{proof}
% Assume that $f\in \Q[X_1,\hdots,X_n]$ has degree at most $D$ and satisfies $\textbf{H}.$ By Theorem 5.1, there exists a Zariski closed set $\D_1\subset \C^{n^2}$ with the property that, if $A \in \C^{n^2}-\D_1$ then, for $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(1), \textbf{H}_i^{'}(2),$ and $\textbf{H}_i^{'}(3).$ Likewise, by Theorem 6.1, we obtain $\D_2 \subset \C^{n^2}$ with the property that if $A \in \C^{n^2}-\D_2$ then $f^A$ satisfies $\textbf{H}_i^{'}(4), \textbf{H}_i^{'}(5),$ and $\textbf{H}_i^{'}(6).$ By Theorem 7.1, $\D_3 \subset \C^{n^2}$ exists with $f^A$ satisfying $\textbf{H}_i^{'}(7)$ when $A \in \C^{n^2}-\D_3$. Clearly then, if 
% \[
% A \in \C^{n^2} - \bigcup_{i=1}^3 \D_i
% \]
% then $f^A$ satisfies  $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$ Take 
% \[
% \Delta = \bigcup_{i=1}^3 \D_i,
% \]
% and note that 
% \[
% \deg \D \leq  \sum_{i=1}^3 \deg \D_i.
% \]
% \end{proof}
% %
% Recall Corollary 3.2: \textit{Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1} D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i \in \{1,\hdots,n\}$ is at least $1-\epsilon.$}
% %
% \begin{proof}
% \[
% Pr[A \in \Delta] \leq \frac{\deg \Delta}{|S|} \leq \frac{D^{\td}}{\epsilon^{-1}D^{\td}} = \epsilon.
% \]
% \end{proof}
% %
% %
% %
% %
\section{Bit complexity}
%Recall Theorem 1.1:   Suppose that $f\in\ZZ[X_1\hdots,X_n]$ is squarefree, satisfies
  $\deg(f) \leq d, \htt(f) \leq s$, and that $V(f) \subset \C^n$ is
  smooth.  Assume also that $f$ is given by a straight-line program
  $\Gamma$ of size $L$ with integer constants of height at most $b.$

@@

  With probability at least $1-\epsilon$, the output is correct. Otherwise, either a proper subset of the points is produced, or FAIL.  In any case, the algorithm uses

@@

\nl Take  $S := \{1,2,\hdots,\epsilon^{-2}3n^2(2d)^{2n}\}$
 and randomly choose $A \in S^{n^2}$, and  $\alpha_{n-1}=(\alpha_1,\hdots,\alpha_{n-1}) \in S^{n-1}$\;

@@

\nl Run \cite[Algorithm 2]{SH} $k \geq -\lg(1-(1+\epsilon)^{(-1/n)})$ times with input 
$\{Y_1-\alpha_1,\hdots,Y_{i-1}-\alpha_{i-1},f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n} \} \textrm{ and } d$\;
\nl Let $\mathscr{Q}_i$ be the highest cardinality zero-dimensional parameterization returned in step 6 \;

@@

    \item We defined $S = \{1,2,\hdots,\epsilon^{-2}3n^2(2d)^{2n}\}$ and therefore the height of $a_{i,j} \in S$ is at most 

@@

\textbf{Probability of success:} Recall Theorem 2.1:   \textit{For $i=1,\dots,n$, there exists a non-zero polynomial $\D_i$ in
  $\C[\A]$ of degree at most $3n(2d)^{2n}$ such that if $\mA \in \C^{n\times
    n}$ does not cancel $\D_i$, then $\mA$ is invertible and $f^\mA$
  satisfies $\textbf{H}_i$ and $\textbf{H}_i^{'}$.} 
 Now, if we set
 \Delta := \bigcup_{i=1}^n \Delta_i,
then
\deg \Delta \leq \sum_{i=1}^n \deg \Delta_i  \leq 3n^2(2d)^{2n},
and for $A \in \C^{n \times n}$ that does not cancel $\Delta,$  $\mA$ is invertible and $f^\mA$
  satisfies $\textbf{H}_i$ and $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$ Let $\mathscr{E}$ denote the event that, with $k$ repetitions of \cite[Algorithm 2]{SH}, for each $i \in \{1,\hdots,n\}$, the result contains all points and is correct.  Then, 

@@

\pr[\textrm{success}] = \pr[A \not\in \Delta] \times \pr[\mathscr{E}].  

@@

By construction of $S$,
\pr[A \not\in \Delta] = 1-\pr[A \in \Delta] \geq
1-\epsilon^2.
Thus, 
\pr[\textrm{success}] &\geq  (1-\epsilon^2)\pr[\mathscr{E}]\\
&\geq (1-\epsilon^2)(1-(11/32)^k)^n\\
& \geq (1-\epsilon^2)(1-2^{-k})^n= 1-\epsilon
if and only if 
(1+\epsilon)(1-2^{-k})^n= 1,
if and only if 
k \geq -\lg\left ( 1 - (1+\epsilon)^{(-1/n)} \right ).
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}
%%
\end{document}
\endinput
%%
%% End of file `sample-xelatex.tex'.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

>>>>

The runtime for cases without a strait-line program in the input are as follows. In these cases a polynomial $f$ of degree $d$ is given in the input, and a strait-line program is built for it by computing all the terms. In this case, $L \geq d^n$ and $b \geq \htt(f) = s,$ so that the number of bit operations becomes
\begin{align*}
&O^{\sim}(nk(d^ns+d^{2n}(s + \log 1/\epsilon^2+d)(d^n+d)\log(s + \log 1/\epsilon^2)))\\
      &= 
O^{\sim}(nk(d^ns+(d^{3n}+d^{2n+1})(s + \log 1/\epsilon^2+d)\log(s + \log 1/\epsilon^2)))\\
     &= 
O^{\sim}(nk(d^{3n}(s + \log 1/\epsilon^2+d)\log(s + \log 1/\epsilon^2)))\\
     &= 
O^{\sim}(nk(d^{3n+1}(s + \log 1/\epsilon^2)\log(s + \log 1/\epsilon^2))),
\end{align*}
and the degree and height of the output does not change.

@@

% \section{Proof of Theorem 2.1}
% %
% Recall Theorem 3.1: \textit{Suppose that $f \in \Q[X_1,\hdots,X_n]$ satisfies \textbf{H}. Then there exists a Zariski closed set $\D \subset \mathbb{C}^{n^2}$ of degree at most $D^{\textrm{\td}}$ such that, if $\textbf{A}$ is chosen from  $\C^{n^2}-\Delta$ ( $\Q^{n^2}-\Delta$?) then $f^A$ satisfies $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$} 
% %
% \begin{proof}
% Assume that $f\in \Q[X_1,\hdots,X_n]$ has degree at most $D$ and satisfies $\textbf{H}.$ By Theorem 5.1, there exists a Zariski closed set $\D_1\subset \C^{n^2}$ with the property that, if $A \in \C^{n^2}-\D_1$ then, for $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(1), \textbf{H}_i^{'}(2),$ and $\textbf{H}_i^{'}(3).$ Likewise, by Theorem 6.1, we obtain $\D_2 \subset \C^{n^2}$ with the property that if $A \in \C^{n^2}-\D_2$ then $f^A$ satisfies $\textbf{H}_i^{'}(4), \textbf{H}_i^{'}(5),$ and $\textbf{H}_i^{'}(6).$ By Theorem 7.1, $\D_3 \subset \C^{n^2}$ exists with $f^A$ satisfying $\textbf{H}_i^{'}(7)$ when $A \in \C^{n^2}-\D_3$. Clearly then, if 
% \[
% A \in \C^{n^2} - \bigcup_{i=1}^3 \D_i
% \]
% then $f^A$ satisfies  $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$ Take 
% \[
% \Delta = \bigcup_{i=1}^3 \D_i,
% \]
% and note that 
% \[
% \deg \D \leq  \sum_{i=1}^3 \deg \D_i.
% \]
% \end{proof}
% %
% Recall Corollary 3.2: \textit{Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1} D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i \in \{1,\hdots,n\}$ is at least $1-\epsilon.$}
% %
% \begin{proof}
% \[
% Pr[A \in \Delta] \leq \frac{\deg \Delta}{|S|} \leq \frac{D^{\td}}{\epsilon^{-1}D^{\td}} = \epsilon.
% \]
% \end{proof}
% %
% %
% %
% %
\section{Bit complexity}
Recall Theorem 1.1:   Suppose that $f\in\ZZ[X_1\hdots,X_n]$ is squarefree, satisfies
  $\deg(f) \leq d, \htt(f) \leq s$, and that $V(f) \subset \C^n$ is
  smooth.  Assume also that $f$ is given by a straight-line program
  $\Gamma$ of size $L$ with integer constants of height at most $b.$

@@

  As part of the randomization, the algorithm chooses a set 
  \[
  S := \{1,2,\hdots,\epsilon^{-2}3n^2(2d)^{3n}\}
  \]
  from which to sample from.

  With probability at least $1-\epsilon$, the output is correct. Otherwise, the algorithm either returns a proper subset of the points or FAIL. In any case, the algorithm uses

@@

\nl Take  $S := \{1,2,\hdots,\epsilon^{-2}3n^2(2d)^{3n}\}$
 and randomly choose $A \in S^{n^2}$, and  $\sigma_{n-1}=(\sigma_1,\hdots,\sigma_{n-1}) \in S^{n-1}$\;

@@

\nl Run \cite[Algorithm 2]{SH} $k \geq -\lg(1-(1+\epsilon)^{(-1/n)})$ times with input (state this in terms of straight line programs?)
$\{Y_1-\sigma_1,\hdots,Y_{i-1}-\sigma_{i-1},f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n} \} \textrm{ and } d$\;
\nl Let $\mathscr{Q}_i$ be the highest cardinality zero-dimensional parameterization returned in step 5 \;

@@

    \item We defined $S = \{1,2,\hdots,\epsilon^{-2}3n^2(2d)^{3n}\}$ and therefore the height of $a_{i,j} \in S$ is at most 

@@

\textbf{Probability of success:} By Theorem 2.1, there exists $\Delta_i \in \C[\A]$ with $\deg \Delta_i \leq 2n^2(2d)^{3n}.$ By Theorem 2.2, there exists $\Xi_i \subset \C[\A]$ with $\deg \Xi_i \leq d^{2n}.$ Take $\Delta =  \prod_{i=1}^n\Delta_i  \Xi_i $ so that 

@@

    \deg \Delta &\leq \sum_{i=1}^n (\deg \Delta_i + \deg \Xi_i)\\
    &\leq 2n^2(2d)^{3n} + nd^{2n}\\
    &\leq 3n^2(2d)^{3n}, 

@@

and note that, for $A \in \C^{n \times n}$ that does not cancel $\Delta,$  $\mA$ is then invertible and $f^\mA$ satisfies $\textbf{H}_i$ and $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$

Let $\mathscr{E}$ be the event that, the parameterizations $[\mathscr{Q}_1,\hdots,\mathscr{Q}_n]$ returned in step 7 of Algorithm 1 contain all points and are correct. Then, 
\begin{align*}
\pr[\textrm{success}] = \pr[A \not\in \Delta] \times \pr[\mathscr{E}~|~A \not\in \Delta].  
\end{align*}
By construction of $S$,
\[
\pr[A \not\in \Delta] = 1-\pr[A \in \Delta] \geq
1-\epsilon^2.
\]
Thus, 
\begin{align*}
\pr[\textrm{success}] &\geq  (1-\epsilon^2)\pr[\mathscr{E}~|~A \not\in \Delta]\\
&\geq (1-\epsilon^2)(1-(11/32)^k)^n\\
& \geq (1-\epsilon^2)(1-2^{-k})^n= 1-\epsilon
\end{align*}
if and only if 
\[
(1+\epsilon)(1-2^{-k})^n= 1,
\]
if and only if 
\[
k \geq -\lg\left ( 1 - (1+\epsilon)^{(-1/n)} \right ).
\]
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}
%%
\end{document}
\endinput
%%
